Individual heterogeneity in the functional topography of the DMN in medial prefrontal cortex 

### 2019-02-08 Manuscript.Rmd
Tweaks to first draft, so everything aligns well

---
title: 'Individual heterogeneity in the functional topography of the DMN in medial prefrontal cortex '
author: "Claudio Toro-Serey, Sean Tobyne, & Joseph T. McGuire"
output:
  pdf_document: default
  word_document: 
    reference_docx: Style_reference.docx
  html_document: default
bibliography: mPFC_paper.bib
header-includes:
    - \usepackage{setspace}\doublespacing
---

@@

## Abstract

@@

The relative high variability seen in the aggregate map could be due to residual noise that was not captured by the sliding window analysis, or variations in the functional organization of mPFC across days. While the sliding window analysis suggested that these partitions are stable, many of the windows contained data from both days. A stricter evaluation of subject-specificity would be to compare every subject's subdivisions produced by data from each day separately. If the variability that we observed was truly due to having captured an individual's unique organization, day to day ARI values should be consistently higher within- than between-subjects. Figure 7 shows pairwise comparisons among ten example subjects for whole brain (left), as well as PCC/mPFC separately (right).

@@

Computing SP on networks from individual sessions shows that we can still coarsely capture a subject's unique organizational pattern (Figure 8), even though the overall levels of agreement decreased (PCC between subjects: mean = `r round(mean(Sess_PCCComparison[[2]]), digits = 2)`, SD = `r round(sd(Sess_PCCComparison[[2]]), digits = 2)`; PCC within subjects: mean = `r round(mean(Sess_PCCComparison[[1]]), digits = 2)`, SD = `r round(sd(Sess_PCCComparison[[1]]), digits = 2)`); mPFC between subjects: mean = `r round(mean(Sess_PFCComparison[[2]]), digits = 2)`, SD = `r round(sd(Sess_PFCComparison[[2]]), digits = 2)`; mPFC within subjects: mean = `r round(mean(Sess_PFCComparison[[1]]), digits = 2)`, SD = `r round(sd(Sess_PFCComparison[[1]]), digits = 2)`). We again computed each subject's ARI ratio in order to quantify the specificity of these partitions, but this time using the mean ARI acros sessions (Figure 9).

@@

To further interrogate the potential effect of noise in our results, we computed the variance explained by activity from each DMN vertex on the mean time series of the DMN community. If the signal quality issues persisted after preprocessing the data, we expected that noisy vertices would have the least explicative power on the activity of the DMN community as a whole. This also implied lower $R^2$ values for mPFC. Indeed, we found that for each subject, $R^2$ values tended to be higher for PCC (`r prcntSig` percent of Wilcoxon ranksum tests across subjects had *p* values < 0.001; mean Cohen's D = `r round(mean(Rsq_PCCvsPFC_ES), digits = 2)`) (Figure 13). These results suggest that, overall, vertices in mPFC tended to be less in synchrony with the activity of the DMN community as a whole. Furthermore, the level of tSNR showed a moderate correlation with $R^2$ values across participants (mean *r* = `r round(mean(noiseCors$R2vtSNR), digits = 2)`; SD = `r round(sd(noiseCors$R2vtSNR), digits = 2)`), meaning that noisier vertices were less in synchrony with the general activity of the DMN.

Given this trend, we asked whether there is a "gradient of noise" along the z-axis of mPFC, such that more ventral vertices would both be noisier and explain the least amount of DMN variance. Here we found that while $R^2$ are somewhat correlated with dorsal-ventral placement across subjects (mean *r* = `r round(mean(noiseCors$ZvR2), digits = 2)`; SD = `r round(sd(noiseCors$ZvR2), digits = 2)`), there is no clear relationship with the z-axis across subjects (Figure 14). Specifically, low $R^2$ are evenly distributed along this axis, denoting a degree of orthogonality between these features. Conversely, tSNR showed a strong linear relationship with the position of the vertex along this axis (mean *r* = `r round(mean(noiseCors$ZvtSNR), digits = 2)`; SD = `r round(sd(noiseCors$ZvtSNR), digits = 2)`). These results contextualize the degree of synchrony of mPFC vertices in terms of potential noise. While still an important factor, it is unlikely that the degree of within-subject pattern stability observed here could be due mostly to highly structured mPFC noise.

@@

We coarsely examined whether these idiosyncratic topographical arrangements were defined by anatomical brain features. To do so, we correlated each subject's Fiedler vector with their respective structural MRI data. On average, we found that the Fiedler vector is uncorrelated with myelin density (mean *r* = `r round(meanMyelin, digits = 2)`; SD = `r round(sdMyelin, digits = 2)`), cortical folding (mean *r* = `r round(meanCurv, digits = 2)`; SD = `r round(sdCurv, digits = 2)`), and surface thickness (mean *r* = `r round(meanThick, digits = 2)`; SD = `r round(sdThick, digits = 2)`). While these findings point at an orthogonality between structure and functional organization, they resonate with previous work that found no strong matches between intersubject functional variability and cortical thickness, and low variability in sulcal depth within DMN [@Mueller2013a].
### *Correlation vs community detection in mPFC*

@@

``` {r Stability per subject: PCC vs mPFC (Fig 5), fig.align="center", fig.width=5, fig.height=3, fig.cap = "MEan agreement between partitions from each sliding window and the map produced with the whole data per subject, for mPFC and PCC separately (left), and the distribution of standard deviations along the time series per subject (right). There was strong (at least 50% after adjusting for chance) and stable agreement in both regions, but higher for PCC.", echo=FALSE, message = FALSE}

>>>>

---
title: 'Individual heterogeneity in the functional topography of the DMN in medial prefrontal cortex '
author: "Claudio Toro-Serey, Sean Tobyne, & Joseph T. McGuire"
output:
  pdf_document: 
    keep_tex: true
  word_document: 
    reference_docx: Style_reference.docx
  html_document: default
bibliography: mPFC_paper.bib
header-includes:
    - \usepackage{setspace}\doublespacing
---

@@

## Abstract (**will be expanded once we decide what to keep**)

@@

The relative high variability seen in the aggregate map could be due to residual noise that was not captured by the sliding window analysis, or variations in the functional organization of mPFC across days. While the sliding window analysis suggested that these partitions are stable, many of the windows contained data from both days. A stricter evaluation of subject-specificity would be to compare every subject's subdivisions produced by data from each day separately. If the variability that we observed was truly due to having captured an individual's unique organization, day to day ARI values should be consistently higher within- than between-subjects. Figure 7 shows pairwise comparisons among ten example subjects for whole brain (left), as well as for PCC and mPFC separately (right).

@@

Computing SP on networks from individual sessions shows that we can still coarsely capture a subject's unique organizational pattern (Figure 9), even though the overall levels of agreement decreased (PCC between subjects: mean = `r round(mean(Sess_PCCComparison[[2]]), digits = 2)`, SD = `r round(sd(Sess_PCCComparison[[2]]), digits = 2)`; PCC within subjects: mean = `r round(mean(Sess_PCCComparison[[1]]), digits = 2)`, SD = `r round(sd(Sess_PCCComparison[[1]]), digits = 2)`); mPFC between subjects: mean = `r round(mean(Sess_PFCComparison[[2]]), digits = 2)`, SD = `r round(sd(Sess_PFCComparison[[2]]), digits = 2)`; mPFC within subjects: mean = `r round(mean(Sess_PFCComparison[[1]]), digits = 2)`, SD = `r round(sd(Sess_PFCComparison[[1]]), digits = 2)`). We again computed each subject's ARI ratio in order to quantify the specificity of these partitions, but this time using the mean ARI acros sessions (Figure 10).

@@

To further interrogate the potential effect of noise in our results, we computed the variance explained by activity from each DMN vertex on the mean time series of the DMN community. If the signal quality issues persisted after preprocessing the data, we expected that noisy vertices would have the least explicative power on the activity of the DMN community as a whole. This also implied lower $R^2$ values for mPFC. Indeed, we found that for each subject, $R^2$ values tended to be higher for PCC (`r prcntSig` percent of Wilcoxon ranksum tests across subjects had *p* values < 0.001; mean Cohen's D = `r round(mean(Rsq_PCCvsPFC_ES), digits = 2)`). These results suggest that, overall, vertices in mPFC tended to be less in synchrony with the activity of the DMN community as a whole. Furthermore, the level of tSNR showed a moderate correlation with $R^2$ values across participants (mean *r* = `r round(mean(noiseCors$R2vtSNR), digits = 2)`; SD = `r round(sd(noiseCors$R2vtSNR), digits = 2)`), meaning that noisier vertices were less in synchrony with the general activity of the DMN.

Given this trend, we asked whether there is a "gradient of noise" along the z-axis of mPFC, such that more ventral vertices would both be noisier and explain the least amount of DMN variance. Here we found that while $R^2$ are somewhat correlated with dorsal-ventral placement across subjects (mean *r* = `r round(mean(noiseCors$ZvR2), digits = 2)`; SD = `r round(sd(noiseCors$ZvR2), digits = 2)`), there is no clear relationship with the z-axis across subjects (Figure 12). Specifically, low $R^2$ are evenly distributed along this axis, denoting a degree of orthogonality between these features. Conversely, tSNR showed a strong linear relationship with the position of the vertex along this axis (mean *r* = `r round(mean(noiseCors$ZvtSNR), digits = 2)`; SD = `r round(sd(noiseCors$ZvtSNR), digits = 2)`). These results contextualize the degree of synchrony of mPFC vertices in terms of potential noise. While still an important factor, it is unlikely that the degree of within-subject pattern stability observed here could be due mostly to highly structured mPFC noise.

@@

We coarsely examined whether these idiosyncratic topographical arrangements were defined by anatomical brain features. To do so, we correlated each subject's Fiedler vector with their respective structural MRI data. On average, we found that the Fiedler vector is uncorrelated with myelin density (mean *r* = `r round(meanMyelin, digits = 2)`; SD = `r round(sdMyelin, digits = 2)`), cortical folding (mean *r* = `r round(meanCurv, digits = 2)`; SD = `r round(sdCurv, digits = 2)`), and surface thickness (mean *r* = `r round(meanThick, digits = 2)`; SD = `r round(sdThick, digits = 2)`). While these findings point at an orthogonality between structure and functional organization, they resonate with previous work that found no strong matches between intersubject functional variability and cortical thickness, and low variability in sulcal depth within DMN [@Mueller2013a].

### *Correlation vs community detection in mPFC*

@@

``` {r Stability per subject: PCC vs mPFC (Fig 5), fig.align="center", fig.width=5, fig.height=3, fig.cap = "Mean agreement between partitions from each sliding window and the map produced with the whole data per subject, for mPFC and PCC separately (left), and the distribution of standard deviations along the time series per subject (right). There was strong (at least 50 percent after adjusting for chance) and stable agreement in both regions, but higher for PCC.", echo=FALSE, message = FALSE}
